---
title: "05_homework_epa_txt_analysis"
author: "Steven Cognac"
date: '2022-04-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(tidyr) #text analysis in R
library(pdftools)
library(lubridate) #working with date data
library(tidyverse)
library(tidytext)
library(readr)
library(quanteda)
library(readtext) #quanteda subpackage for reading pdf
library(quanteda.textstats)
library(quanteda.textplots)
library(ggplot2)
library(forcats)
library(stringr)
library(quanteda.textplots)
library(widyr)# pairwise correlations
library(igraph) #network plots
library(ggraph)
library(here)
```


# Import data

```{r pdf_import}
files <- list.files(path = here("05_EPA_txt_analysis"), pattern = "pdf$", full.names = T)

ej_reports <- lapply(files, pdf_text)

ej_pdf <- readtext("*.pdf", docvarsfrom = "filenames", 
                    docvarnames = c("type", "year"),
                    sep = "_")

#creating an initial corpus containing our data
epa_corp <- corpus(x = ej_pdf, text_field = "text" )
summary(epa_corp)

#I'm adding some additional, context-specific stop words to stop word lexicon
more_stops <-c("2015","2016", "2017", "2018", "2019", "2020", "www.epa.gov", "https")
add_stops<- tibble(word = c(stop_words$word, more_stops)) 
stop_vec <- as_vector(add_stops)

```

1.  What are the most frequent trigrams in the dataset? How does this compare to the most frequent bigrams? Which n-gram seems more informative here, and why?

```{r quanteda_init}
# output is a list of vectors
tokens <- tokens(epa_corp, remove_punct = TRUE)

# some token cleaning
toks1 <- tokens_select(tokens, min_nchar = 3)
toks1 <- tokens_tolower(toks1)
toks1 <- tokens_remove(toks1, pattern = (stop_vec))

# two words is a "bi-gram" "two-word pairs" as a fundamental unit of analysis
toks2 <- tokens_ngrams(toks1, n=2)
toks3 <- tokens_ngrams(toks1, n=3)

# create document feature matrix from tokens
# rows refer to number of occurrences within entire corpus in entire document
dfm2 <- dfm(toks2)
dfm2 <- dfm_remove(dfm2, pattern = c(stop_vec))

dfm3 <- dfm(toks3)
dfm3 <- dfm_remove(dfm3, pattern = c(stop_vec))

# calculate basic statistics
freq_words2 <- textstat_frequency(dfm2, n=20, groups = year)
freq_words2$token <- rep("bigram", 20)

freq_words3 <- textstat_frequency(dfm3, n=20, groups = year)
freq_words3$token <- rep("trigram", 20)
```

```{r}
freq_compare <- rbind(freq_words3 %>% slice_max(frequency, n = 10),
                      freq_words2 %>% slice_max(frequency, n = 10)) %>% 
  mutate(order = reorder_within(feature, frequency, token))

freq_compare %>% 
  ggplot(aes(y = reorder(feature, desc(frequency)), x = frequency), fill = token) +
  geom_col() +
  facet_wrap(~token, scales = "free") +
  labs(y = "Feature Words",
       x = "Word Frequency",
       title = "Top 10 Bigram and Trigram Words",
       subtitle = "2015-2020 EPA Annual Environmental Justice Progress Reports") +
  scale_y_discrete(limits=rev)
```


2.  Choose a new focal term to replace "justice" and recreate the correlation table and network (see corr_paragraphs and corr_network chunks). Explore some of the plotting parameters in the cor_network chunk to see if you can improve the clarity or amount of information your plot conveys. Make sure to use a different color for the ties!

```{r tidy}
#convert to tidy format and apply my stop words
raw_text <- tidy(epa_corp)

#Distribution of most frequent words across documents
raw_words <- raw_text %>%
  mutate(year = as.factor(year)) %>%
  unnest_tokens(word, text) %>%
  anti_join(add_stops, by = 'word') %>%
  count(year, word, sort = TRUE)

#number of total words by document  
total_words <- raw_words %>% 
  group_by(year) %>% 
  summarize(total = sum(n))

report_words <- left_join(raw_words, total_words)

# tokenizing to the "paragraph" level
# helps extract meaning from words by attaching a paragraph ID
par_tokens <- unnest_tokens(raw_text, output = paragraphs, input = text, token = "paragraphs")

par_tokens <- par_tokens %>%
 mutate(par_id = 1:n())

par_words <- unnest_tokens(par_tokens, output = word, input = paragraphs, token = "words")

```


```{r corr_paragraphs}
# create correlations between words
word_cors <- par_words %>% 
  add_count(par_id) %>% 
  filter(n >= 50) %>% 
  select(-n) %>%
  pairwise_cor(word, par_id, sort = TRUE)

# select correlations that contain "justice"
fuels_cors <- word_cors %>% 
  filter(item1 == "fuels")

# filtering words in the context of `item1`
word_cors %>%
  filter(item1 %in% c("gasoline", "energy", "coal", "wind"))%>%
  group_by(item1) %>%
  top_n(5) %>%
  ungroup() %>%
  mutate(item1 = as.factor(item1),
  name = reorder_within(item2, correlation, item1)) %>%
  ggplot(aes(y = name, x = correlation, fill = item1)) + 
  geom_col(show.legend = FALSE) +
  facet_wrap(~item1, ncol = 2, scales = "free")+
  scale_y_reordered() +
  labs(y = NULL,
         x = NULL,
         title = "Correlations with key words",
         subtitle = "EPA EJ Reports")
  
  #let's zoom in on just one of our key terms
fuels_cors <- word_cors %>%
  filter(item1 == "fuels") %>%
  mutate(n = 1:n())
 
```


```{r corr_network}
fuels_cors  %>%
  filter(n <= 40) %>%
  graph_from_data_frame() %>%
  ggraph(layout = "fr") +
  geom_edge_link(aes(edge_alpha = correlation, edge_width = correlation), edge_colour = "firebrick3") +
  geom_node_point(size = 4) +
  geom_node_text(aes(label = name), repel = TRUE, 
                 point.padding = unit(0.2, "lines")) +
  theme_void()
```


3.  Write a function that allows you to conduct a keyness analysis to compare two individual EPA reports (hint: that means target and reference need to both be individual reports). Run the function on 3 pairs of reports, generating 3 keyness plots.

```{r}
# target = 1 means first document
keyness <- textstat_keyness(dfm2, target = 2)

# Interpreting graph, in 2015 document the word *climate* occured more than one would expect given the reference documents (remaining EPA PDF documents 2016-2020)
textplot_keyness(keyness)
```

```{r}
for (i in length()unique(ej_pdf$doc_id))
```


```{r}
scores <- function(feature_matrix) {
  
  for (f in seq_along(ej_pdf$doc_id)){
    
    keyness <- textstat_keyness(feature_matrix, [f])
    
    plot <- textplot_keyness(keyness)
  }
  
}
```





4.  Select a word or multi-word term of interest and identify words related to it using windowing and keyness comparison. To do this you will create to objects: one containing all words occurring within a 10-word window of your term of interest, and the second object containing all other words. Then run a keyness comparison on these objects. Which one is the target, and which the reference? [Hint](https://tutorials.quanteda.io/advanced-operations/target-word-collocations/)

```{r}

```

