---
title: 'Topic 7: Word Embeddings'
author: "Steven Cognac"
date: '2022-05-11'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r packages, include = FALSE}
library(here)
library(tidytext)
library(tidyverse)
library(widyr)
library(irlba) #singluar value decomposition
library(broom) # creating search_synonym function
library(textdata)
library(ggplot2)
library(dplyr)
library(here)
library(readr)
```

# What are word Embeddings?

Word embeddings are a type of word representation that allows words with similar meaning to have a similar representation. In Natural Language Processing for text analysis, they typically come in the form of real-valued vectors that encode the meaning of words such that the words closer together are expected to be similar in meaning.

For this exercise, we'll use pre-trained word vector from the [GloVe: Global Vectors for Word Representation](https://nlp.stanford.edu/projects/glove/) project. Specifically, we'll use the Wiki2014 + Gigword5 [300d vector file](https://nlp.stanford.edu/data/glove.6B.zip).

## Import data
I downloaded the data and saved it to a local folder

```{r, results='hide'}
glov_df <- read_table(here('07_word_embeddings/data/glove.6B.300d.txt'), 
                       col_names = FALSE) %>% 
  column_to_rownames(., var = "X1")
```

## Create similarity function

```{r syn-function}
search_synonyms <- function(glov_df, selected_vector) {

  dat <- glov_df %*% selected_vector
  
  similarities <- dat %>%
        tibble(token = rownames(dat), 
               similarity = dat[,1])
  
  similarities %>%
       arrange(-similarity) %>%
        select(c(2,3))
}
```

## Word Synonyms

First we'll convert the dataframe to a matrix
```{r}
glov_matrix <- data.matrix(glov_df)
```

You can search word synonyms with the `search_synonyms()` function. We can then plot those synonyms.

```{r find-synonyms}
king <- search_synonyms(glov_matrix, glov_matrix["king",])
man <- search_synonyms(glov_matrix, glov_matrix["man",])
```


```{r plot-synonyms}
king_man_plot <- king %>%
  mutate(selected = "king") %>%
  bind_rows(man %>%
              mutate(selected = "man")) %>%
  group_by(selected) %>%
  top_n(15, similarity) %>%
  ungroup %>%
  mutate(token = reorder(token, similarity)) %>%
  
  # plot setup
  ggplot(aes(token, similarity, fill = selected)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~selected, scales = "free") +
  coord_flip() +
  theme(strip.text=element_text(hjust=0, size=12)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = NULL, 
       title = "What word vectors are most similar to king or man?",
       subtitle = "Top 15 words")

king_man_plot
```

In our word embedding plots, we see `king` is highly similar to other terms of royalty. These are words that have a high similarity to and in some cases, be swapped for that word. The word `man` on the other hand is more similar to other words that relate to people.

## Word Math

You can use simple arithmetic to search for synonyms of multiple words "king + man" or synonyms of one word while excluding the other "king - man".

Here we'll try a couple examples
```{r word-math}
king_no_man <- glov_matrix["king",] - glov_matrix["man",] 
search_synonyms(glov_matrix, king_no_man)

# love and sad
love <- search_synonyms(glov_matrix, glov_matrix["love",])
sad <- search_synonyms(glov_matrix, glov_matrix["sad",])


love_no_sad <- glov_matrix["love",] - glov_matrix["sad",]
search_synonyms(glov_matrix, love_no_sad)

love_sad <- glov_matrix["love",] + glov_matrix["sad",]
search_synonyms(glov_matrix, love_sad)


# alcohol and drug
alcohol <- search_synonyms(glov_matrix, glov_matrix["alcohol",])
drug <- search_synonyms(glov_matrix, glov_matrix["drug",])

alcohol_no_drug <- glov_matrix["alcohol",] - glov_matrix["drug",]
alcohol_no_drug <- search_synonyms(glov_matrix, alcohol_no_drug)

alcohol_drug <- glov_matrix["alcohol",] + glov_matrix["drug",]
alcohol_drug <- search_synonyms(glov_matrix, alcohol_drug)
```

```{r}
alcohol_drug_plot <- alcohol_no_drug %>%
  mutate(selected = "04_alcohol-drug") %>%
  bind_rows(alcohol %>% 
              mutate(selected = "01_alcohol"),
            drug %>% 
              mutate(selected = "03_drug"),
            alcohol_drug %>%
              mutate(selected = "02_alcohol+drug")) %>%
  group_by(selected) %>%
  top_n(15, similarity) %>%
  ungroup %>%
  mutate(token = reorder(token, similarity)) %>%
  
  # plot setup
  ggplot(aes(token, similarity, fill = selected)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~selected, scales = "free") +
  coord_flip() +
  theme(strip.text=element_text(hjust=0, size=12)) +
  scale_y_continuous(expand = c(0,0)) +
  labs(x = NULL, 
       title = "Word vectors for alcohol & drug combinations",
       subtitle = "Top 15 words")

alcohol_drug_plot
```